\textbf{Proof Reconstruction in Classical Propositional Logic}

### 2017-04-24 slides/slides.tex
[ slides ] clickeable items in the diagram.

\documentclass[xetex,hyperref={pdfpagelabels=false}]{beamer}

@@

% \setbeamercolor{institute}{fg=blu}
% \setbeamercolor{subtitle}{fg=blu}
% \setbeamercolor{titlelike}{fg=blu}
\setbeamerfont{footnote}{size=\tiny}
\setbeamercolor{footnote}{fg=gray}
% \setbeamercolor{block title}{bg=blue,fg=blu}
% \setbeamercolor{block body}{bg=aliceblue}
% \setbeamercolor{item}{fg=blu} % color of bullets
% \setbeamercolor{subitem}{fg=blu}
% \setbeamercolor{itemize/enumerate subbody}{fg=blu}
% \setbeamertemplate{itemize subitem}{{\textendash}}
% \setbeamerfont{itemize/enumerate subbody}{size=\footnotesize}
% \setbeamerfont{itemize/enumerate subitem}{size=\footnotesize}

@@

\usepackage[english]{babel}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{bussproofs}
\usepackage{fancyvrb}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{lstagda}
\usepackage{tikz}
\usepackage{url}

@@

\begin{document}

\setcounter{page}{1}

\begin{frame}[plain]
% \tikz[overlay,remember picture] \node[opacity=0.05, at=(6,6)] {
% $Γ ⊢ ϕ$
% };
\titlepage
\end{frame}

@@

\section{Introduction}
\subsection{Motivation}

@@

\begin{frame}[fragile]{Translation}

\only<1-2>{At the moment, the communication between Agda and
the ATPs is unidirectional because the ATPs are being used as oracles.\citep{Sicard2015}.}
\only<1-2>{\vfill}

\begin{tikzpicture}[scale=2]
\only<1->{ \node(agda-pragmas) at (0,0){\includegraphics{figures/agda-pragmas.pdf}}};
\only<1>{
  \node[fill=aliceblue, rectangle](atp-pragma) at (3,0){
  \scalebox{0.7}{\usebox\agdapragma}
  }};

\only<2->{ \node(eagda) at (2,0){
  \includegraphics{figures/eagda.pdf}
  \footnote{Development version of Agda in order to handle a new built-in ATP-pragma.
    \url{https://github.com/asr/eagda}}
  }};

\only<3->{ \node(apia) at (4,0){
  \includegraphics{figures/apia.pdf}
  \footnote{Haskell program for proving first-order theorems written in Agda using ATPs.
    \url{https://github.com/asr/apia}}}
  };

\only<3->{ \node(tptp) at (2,-2){\includegraphics{figures/tptp.pdf}}};

\only<3->{ \node(atp) at (4,-2){\includegraphics{figures/atp.pdf}}};

@@

\subsection{Automatic Provers}

\subsubsection{TPTP Syntax}

\begin{frame}[fragile]{TPTP Syntax}{Thousands of Problems for Theorem Provers}
\label{tptp-syntax}
 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 0cm)
    {\includegraphics[width=0.15\textwidth]{figures/tptp}};

  \begin{itemize}
  \item Is a language\footnote{Is available at
        \url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}}
        to encode problems in text files
  \item Is the input of the ATPs
  \item his problems contains formulas with the form
   \begin{center}
\begin{verbatim}
language(name, role, formula).
\end{verbatim}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] THF, TFF, FOF, or CNF
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture, among others
      \item[\texttt{formula}] the logic formula in the language
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{TPTP Examples}{Prop-Pack Repository}
\label{tptp-examples}
Problems in Propositional Logic:
  \begin{itemize}
    \item $p ⊢ p$
\begin{tptp}
$ cat basic-4.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
\end{tptp}
    \item $p ∧ q ⊢ q ∧ p$
\begin{tptp}
$ cat conj-3.tptp
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
\end{tptp}
    \item $⊢ ¬ (p ∧ ¬ p) ∨ (q ∧ ¬ q)$
\begin{tptp}
$ cat neg-7.tptp
fof(goal, conjecture, ~ ((p & ~ p) | (q & ~ q))).
\end{tptp}
    \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{TSTP Syntax}
\label{tstp-syntax}
 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, -0.5cm)
    {\includegraphics[width=0.15\textwidth]{figures/tstp}};

A TSTP derivation
\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/QuickGuide/Derivations.html}}
\begin{itemize}
  \item Is a Directed Acyclic Graph where
  \begin{itemize}
    \item[\texttt{leaf}] is a formulae from the TPTP input
    \item[\texttt{node}] is a formulae inferred from parent formulae
    \item[\texttt{root}] the final derived formulae
  \end{itemize}
  \item Is a list of annotated formulae:
  {\footnotesize
    \begin{center}
\begin{verbatim}
language(name,role,formula,source[,useful info]).
\end{verbatim}
  \end{center}}
\begin{itemize}
  \item[\texttt{source}] typically is a inference record:
    {\footnotesize
\begin{verbatim}
  inference(rule, [], [])
  \end{verbatim}}
\begin{itemize}
    \item[-] The inference rule name,
    \item[-] a list of useful inference information, and
    \item[-] a list of references to its parent formulae.
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{TSTP Example}
\label{tstp-example}
 \tikz[overlay,remember picture]
  \node at (0.93\textwidth, -0.3cm)
    {\includegraphics[width=0.15\textwidth]{figures/metis}};

@@

\begin{frame}[fragile]{DAG for the previous TSTP derivation found by \texttt{Metis} ATP}
\label{tstp-dag}
\vfill
\begin{center}
\includegraphics[height=0.8\textheight]{figures/derivation.png}
\end{center}
\end{frame}

\begin{frame}{Design decisions for the Reconstruction Tool}
\label{agda}

 \tikz[overlay,remember picture]
  \node at (0.93\textwidth, 0cm)
    {\includegraphics[width=0.2\textwidth]{figures/athena}};

\textbf{Programming Languages}
\begin{itemize}
\item \texttt{Haskell} is a standardized, general-purpose purely functional programming language.\\
Our usages:
\begin{itemize}
    \item Parsing
    \item AST construction
    \item Creation and analysis of DAG derivations
    \item Analysis of inference rules used
    \item Generation of Agda code of the proof
\end{itemize}

\item \texttt{Agda} is dependently typed functional programming language and it also a proof assistant.\\
Our usages:
    \begin{itemize}
    \item Logic framework for Classical Propositional Logic
    \item Type-Checker for the Agda code of the proofs generated
    \end{itemize}
\end{itemize}

@@

\begin{frame}{References}
\label{references}
\printbibliography
\end{frame}

>>>>

\documentclass[xetex, hyperref={pdfpagelabels=false}]{beamer}

@@

% \setbeamercolor{institute}{fg=blu}
\setbeamercolor{title}{fg=plum}
% \setbeamercolor{subtitle}{fg=blu}

% \setbeamercolor{titlelike}{fg=blu}
\setbeamerfont{footnote}{size=\tiny}
\setbeamercolor{footnote}{fg=gray}
% \setbeamercolor{block title}{bg=blue,fg=blu}
% \setbeamercolor{block body}{bg=aliceblue}
% \setbeamercolor{item}{fg=blu} % color of bullets
% \setbeamercolor{subitem}{fg=blu}
% \setbeamercolor{itemize/enumerate subbody}{fg=blu}
% \setbeamertemplate{itemize subitem}{{\textendash}}
% \setbeamerfont{itemize/enumerate subbody}{size=\footnotesize}
% \setbeamerfont{itemize/enumerate subitem}{size=\footnotesize}

@@

\usepackage[english]{babel}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{bussproofs}
\usepackage{fancyvrb}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{mylst}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}

\usepackage{url}

@@

\begin{document}
\setcounter{page}{1}

\tikzset{
    hyperlink node/.style={
        alias=sourcenode,
        append after command={
            let     \p1 = (sourcenode.north west),
                \p2=(sourcenode.south east),
                \n1={\x2-\x1},
                \n2={\y1-\y2} in
            node [inner sep=0pt, outer sep=0pt,anchor=north west,at=(\p1)] {\hyperlink{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
                    %xelatex needs \XeTeXLinkBox, won't create a link unless it
                    %finds text --- rules don't work without \XeTeXLinkBox.
                    %Still builds correctly with pdflatex and lualatex
        }
    }
}

\begin{frame}
\titlepage
\begin{tikzpicture}[overlay,remember picture, scale=0.5]
\foreach \x in {0,4,..., \textwidth}
  \foreach \y in {0,1,..., 20} {
    \node[opacity=0.05](tptp) at (\x,\y){$Γ ⊢ ϕ$};
  }
\end{tikzpicture}
\end{frame}

@@

\section{Introduction}

\begin{frame}[fragile, label=reconstruction]{Proof Reconstruction}

\begin{tikzpicture}
\node(problem) at (0,0)
   {\includegraphics[scale=0.7]{figures/problem}};

\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};

\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}};

\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};

\node[below= 2cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}};

\node[left = 0.5cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}};

\node[left = 0.5cm of agdafile, hyperlink node=agda] (agda)
  {\includegraphics[scale=0.7]{figures/agda}};

\node[below = 0.2cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};

\node[below = 0.2cm of verified, opacity=0.5] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};

\draw[->, thick] (problem) to (tptp);
\draw[->, thick] (tptp) to (metis);
\draw[->, thick] (metis) to (tstp);
\draw[->, thick] (tstp) to (athena);
\draw[->, thick] (athena) to (agdafile);
\draw[->, thick] (agdafile) to (agda);
\draw[->, thick] (agda) to (verified);
\draw[->, thick, gray] (agda) to (failure);
\end{tikzpicture}
\vfill
\end{frame}

\subsection{Motivation}

@@

\begin{frame}[fragile]{Previous Work: Apia}{Proving first-order theorems written in Agda using automatic theorem provers for first-order logic}

\only<1-2>{At the moment, the communication between Agda and
the ATPs is unidirectional because the ATPs are being used as oracles \citep{Sicard2015}.}
\only<1-2>{\vfill}

\begin{tikzpicture}[scale=2]
\only<1->{ \node(agda-pragmas) at (0,0){\includegraphics{figures/agda-pragmas}}};
\only<1>{
  \node[fill=aliceblue, rectangle](atp-pragma) at (3,0){
  \scalebox{0.7}{\usebox\agdapragma}
  }};

\only<2->{ \node(eagda) at (2,0){
  \includegraphics{figures/eagda}
  \footnote{Development version of Agda in order to handle a new built-in ATP-pragma.
    \url{https://github.com/asr/eagda}}
  }};

\only<3->{ \node(apia) at (4,0){
  \includegraphics{figures/apia}
  \footnote{Haskell program for proving first-order theorems written in Agda using ATPs.
    \url{https://github.com/asr/apia}}}
  };

\only<3->{ \node(tptp) at (2,-2){\includegraphics{figures/tptp}}};

\only<3->{ \node(atp) at (4,-2){\includegraphics{figures/atp}}};

@@

% \subsection{Automatic Provers}

% \subsubsection{TPTP Syntax}

\begin{frame}{}
{\Huge Bonus Slides}
\end{frame}

\begin{frame}[fragile, label=tptp-syntax]{TPTP Syntax}{Thousands of Problems for Theorem Provers}

 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 0cm)
    {\includegraphics[width=0.15\textwidth]{figures/tptp}};

  \begin{itemize}
  \item Is a language\footnote{Is available at
        \url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}}
        to encode problems in text files
  \item Is the input of the ATPs
  \item his problems contains formulas with the form
   \begin{center}
\begin{verbatim}
language(name, role, formula).
\end{verbatim}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF, or CNF %THF, TFF,
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture, among others
      \item[\texttt{formula}] the logic formula in the language
    \end{itemize}
  \end{itemize}

\hyperlink{reconstruction}{Go to Reconstruction}
\end{frame}

\begin{frame}[fragile]{TPTP Examples}{From Prop-Pack repository\footnote{Is available at \url{http://github.com/jonaprieto/pro-pack}}}
\label{tptp-examples}
Problems in Propositional Logic:
  \begin{itemize}
    \item $p ⊢ p$
\begin{tptp}
$ cat basic-4.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
\end{tptp}
    \item $p ∧ q ⊢ q ∧ p$
\begin{tptp}
$ cat conj-3.tptp
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
\end{tptp}
    \item $⊢ ¬ (p ∧ ¬ p) ∨ (q ∧ ¬ q)$
\begin{tptp}
$ cat neg-7.tptp
fof(goal, conjecture, ~ ((p & ~ p) | (q & ~ q))).
\end{tptp}
    \end{itemize}
\end{frame}

@@

\begin{frame}[fragile, label=tstp-syntax]{TSTP Syntax}

 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, -0.5cm)
    {\includegraphics[width=0.15\textwidth]{figures/tstp}};

A TSTP derivation
\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/QuickGuide/Derivations.html}}
\begin{itemize}
  \item Is a Directed Acyclic Graph where
  \begin{itemize}
    \item[\texttt{leaf}] is a formulae from the TPTP input
    \item[\texttt{node}] is a formulae inferred from parent formulae
    \item[\texttt{root}] the final derived formulae
  \end{itemize}
  \item Is a list of annotated formulae:
  \end{itemize}

\begin{center}
  {\footnotesize
\begin{verbatim}
language(name, role, formula, source [,useful info]).
\end{verbatim}}
\end{center}

where \textbf{source} typically is a inference record:
\begin{center}
{\footnotesize
\begin{verbatim}
inference(rule, useful information, parents)
\end{verbatim}
}
\end{center}

\begin{frame}[fragile, label=tstp-example]{TSTP Example}

 \tikz[overlay,remember picture]
  \node at (0.93\textwidth, -0.3cm)
    {\includegraphics[width=0.15\textwidth]{figures/metis}};

@@

\begin{frame}[fragile, label=tstp-dag]{DAG for the previous TSTP derivation found by \texttt{Metis} ATP}

\vfill
\begin{center}
\includegraphics[height=0.8\textheight]{figures/derivation.png}
\end{center}
\end{frame}

\begin{frame}[label=haskell]{Design Decisions for the Reconstruction Tool}
 \tikz[overlay,remember picture]
  \node at (0.93\textwidth, 0cm)
    {\includegraphics[width=0.2\textwidth]{figures/athena}};

\texttt{Haskell} % is a standardized, general-purpose purely functional programming language.\\ Our usages: 
\begin{itemize}
    \item Parsing
    \item AST construction
    \item Creation and analysis of DAG derivations
    \item Analysis of inference rules used
    \item Generation of Agda code of the proof
\end{itemize}
\end{frame}

\begin{frame}[label=agda]{Design Decisions for the Reconstruction Tool}{Programming Languages}
\texttt{Agda} is a dependently typed functional programming language and it also a proof assistant.\\
Our usages:
    \begin{itemize}
    \item Logic framework for Classical Propositional Logic
    \item Type-Checker for the Agda code of the proofs generated
    \end{itemize}

@@

\begin{frame}[label=references]{References}
\label{references}
\printbibliography
\end{frame}
